---
title: "The Investigation of the Exponential Distribution"
author: "Tianxing Li"
date: "Friday, February 20, 2015"
output: pdf_document
---

## Overview

This is the project for the statistical inference class. This report investigated the exponential distribution in R and compared it with the Central Limit Theorem. 


## Simulations

1. The simulation runs on the following parameters.
```{r, echo=TRUE}
loop <- 1000
lambda <- 0.2
n <- 40
seed <- 100
```

2. Run the simulation and format the result to a matrix.
```{r, echo=TRUE}
set.seed(seed)
rawResult <- rexp(loop * n, lambda)
result <- matrix(rawResult, loop)
```

3. Calculate the sample mean and variance.
```{r, echo=TRUE}
sampleMean <- rowMeans(result)
sampleVar <- apply(result, 1, var)
```


## Sample Mean versus Theoretical Mean

Plot the histogram of the sample mean as below.
```{r, echo=TRUE, message=FALSE, fig.height=3}
#hist(sampleMean)
library(ggplot2)
m <- qplot(sampleMean, geom="histogram", main="Histogram of Sample Mean", 
           xlab="mean", ylab="count")
m + geom_histogram(aes(fill = ..count..))
```

In the simulation, we set $\lambda$ to `r lambda`. The mean of exponential distribution is 1/$\lambda$, which is `r 1 / lambda`. The simulated data sample has values for mean of `r mean(sampleMean)`, which is close to the expected value.


## Sample Variance versus Theoretical Variance

In the simulation, we set $\lambda$ to `r lambda`. The mean of exponential distribution is (1/$\lambda$)^2, which is `r (1 / lambda)^2`. Plot the histogram of the sample variance as below, as well as the difference between the sample Variance and theoretical Variance, which seems to obey gamma distribution.
```{r, echo=TRUE, message=FALSE, fig.height=3}
library(gridExtra)
m1 <- qplot(sampleVar, geom="histogram", main="Histogram of Sample Variance", 
           xlab="variances", ylab="count")
m1 <- m1 + geom_histogram(aes(fill = ..count..))

diff <- sampleVar - (1 / lambda)^2
m2 <- qplot(diff, geom="histogram", main="Difference of the Variances", 
           xlab="variances", ylab="count")
m2 <- m2 + geom_histogram(aes(fill = ..count..))
grid.arrange(m1, m2, ncol=2)
```


## Distribution

We can tell the distribution is approximately normal by examining the empirical cumulative distribution and Quantile-Quantile (Q-Q) plot. R allows to compute the empirical cumulative distribution function by `ecdf()` and R also provides `qqnorm()` to get Quantile-Quantile (Q-Q) plot in order totest the goodness of fit of a gaussian distribution.
```{r, echo=TRUE, message=FALSE, fig.height=3}
defaultPar <- par(mfrow=c(1, 2))
plot(ecdf(rawResult), main="Empirical Cumulative Distribution")
qqnorm(sampleMean)
qqline(sampleMean)
par(defaultPar)
```
